\LARGE Untangling the link between traits, size and growth rate in plants

### 2016-07-29 ms/ms.tex
Outline of main structure, embed eqn references in table and figure

@@

% Enable cross referencing to Evidence document
\usepackage{xr}
\externaldocument{ms-supp}

@@

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item We ...
\end{enumerate}

\section{Ideas}\label{abstract}

New refs: \citep{Ishihara-2016, Duursma-2016, Gibert-2016, Falster-2016, Kunstler-2015,Falster-2015b, Falster-2015a,Wenk-2015,Li-2014}

Table near intsro summarising empirical patterns to be explained by integrated model?

\section{Abstract}\label{abstract}

\section{Introduction}\label{introduction}

@@

\section{Materials and Methods}\label{materials-and-methods}

\subsection{Growth data}\label{growth-data}

\subsection{Conceptual framework}\label{conceptual-framework}

@@

Eqs. \ref{eq:dhdt}-\ref{eq:dDdt} are mathematically true, and must
therefore hold for any model of plant growth. To make explicit
predictions, however, requires two additional elements: 1) A specific
model describing how the various mass and area terms
($M_{\rm l}, M_\textrm{s}, M_{\rm b}, M_{\rm r}, A_{\rm l}, A_\textrm{s}, A_{\rm b}$)
vary relative to one another; and 2) For any trade-offs in plant
function to be specified; this is where traits have an influence.  These two components are outlined in following sections.

\subsubsection{Standard model for mass production}

@@

\begin{equation}\label{eq:dbdt}
\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
  = \underbrace{\strut y}_\textrm{yield}
    \big( \underbrace{\strut A_{\rm l} \, p}_\textrm{photosynthesis} -
     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
\end{equation}

Here, $m,r$, and $k$ refer to the mass, respiration rate, and
turnover rate of different tissues, denoted by subscripts $l$=leaves,
$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
assimilated carbon fixed in biomass (the remaining fraction being lost
as growth respiration) (see Table \ref{tab:params} for units and
definitions). Photosynthesis is proportional to leaf area,
$A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per area.
The total mass of living tissues $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_\textrm{s}+M_{\rm r}.$

The only potentially controversial element of in eq. \ref{eq:dbdt} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}. This assumption is likely to hold true for most woody plants, because the ..... However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegtative part of the plant (refs XX) \citep{ Thornley-2000}.

\subsubsection{Decomposing height, basal area and diameter growth}

To model growth in either height, basal area or stem diameter requires that account not just for mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout. Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{Falster-2011}:

\begin{equation} \label{eq:dhdt}
\underbrace{\strut\frac{{\rm d}H}{{\rm d}t}}_\textrm{height growth rate}= \underbrace{\strut\frac{{\rm d}H}{{\rm d}A_{\rm l}}}_\textrm{architecture}
\times \underbrace{\strut\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}}_\textrm{marginal leaf deployment}
\times \underbrace{\strut\frac{{\rm d}M_{\rm a}}{{\rm d}B}}_\textrm{allocation to growth}
\times \underbrace{\strut\frac{{\rm d}B}{{\rm d}t}}_\textrm{mass production}.
\end{equation}

@@

\begin{equation}\label{eq:daldmt}
\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}}
= \big(\phi
 + \frac{{\rm d}M_\textrm{s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}\big)^{-1}.
\end{equation}

@@

In a similar way, basal area ($A_{\rm st}$) increment can be
expressed as the sum of increments in sapwood, bark \& heartwood areas
($A_\textrm{ss}, A_{\rm sb}, A_{\rm sh}$ respectively):
$\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm sb}}{{\rm d}t} + \frac{{\rm d}A_\textrm{ss}}{{\rm d}t} + \frac{{\rm d}A_{\rm sh}}{{\rm d}t}$.
We now have an equation for basal area growth that conatins many of the same
elements as eq. \ref{eq:dhdt}:

\begin{equation}\label{eq:dast}
\frac{{\rm d}A_{\rm st}}{{\rm d}t}=
\left(\frac{{\rm d}A_\textrm{ss}}{{\rm d}A_{\rm l}} + \frac{{\rm d}A_{\rm sb}}{{\rm d}A_{\rm l}}\right) \times
\frac{{\rm d}A_{\rm l}}{{\rm d}M_{\rm a}} \times \frac{{\rm d}M_{\rm a}}{{\rm d}B} \times \frac{{\rm d}B}{{\rm d}t} +\frac{{\rm d}A_{\rm sh}}{{\rm d}t} .
\end{equation}

Diameter growth is then given by the geometric relationship between stem
diameter ($D$) and $A_{\rm st}$:

\begin{equation} \label{eq:dDdt}
\frac{{\rm d}D}{{\rm d}t}= \sqrt{\frac{\pi}{A_{\rm st}}} \frac{{\rm d}A_{\rm st}}{{\rm d}t}.
\end{equation}

@@

To address the first of these challenges, we use a functional balance
model \citep{Yokozawa-1995, Falster-2011} assuming
1) an allometric relation between total leaf area and plant height, and
2) a constant ratio between total leaf area and each of sapwood
cross-sectional area, bark cross-sectional area, and total root mass.
These assumptions, which are well-supported by data (Fig.
\ref{f-assumptions}), lead immediately to an allometric model for plant
size (Table \ref{tab:allometry}, see Supplementary material for
derivations). Substituting from Table \ref{tab:allometry} into eqs.
\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
for plant growth in relation to size.

@@

\subsubsection{Embeeding trait-related trade-offs}

@@

As for leaf, cheaper stem construction (lower $\rho$) decreases the
cost of deploying a unit of leaf area, and may thereby increase growth
rates. However, the cost of cheap stem construction must also be
accounted for. However, the direct physiological trade-offs of lower
$\rho$ are less understood than for $\phi$. One possibility is that
lower $\rho$ decreases mechanical strength resulting in higher
mortality \citep{Chave-2009, Wright-2010}. Under
this scenario, decreasing $\rho$ will always provide a growth
advantage, because the costs of low $\rho$ are not realised within the
terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
\ref{tab:trade-offs}). In addition, lower $\rho$ may increase turnover
of sapwood. Mirroring our assumptions for leaves, we let sapwood
turnover decrease with $\rho$,
i.e.$k_\textrm{ss}=\alpha_5 \, \rho^{-\beta_5}$, with $\beta_5 > 1$.

@@

\subsection{Analysis}\label{analysis}

\newpage
\section{Results and Discussion}\label{results}

@@

This new growth model captures the intrinsically size-dependent nature
of plant growth, recovering well-known
empirical patterns \citep{Sillett-2010, King-2011}. Height growth shows a hump-shaped pattern
with size, first increasing then decreasing. This pattern results from
systematic changes in the four components of eq. \ref{eq:dhdt} with size
(Fig. \ref{f-conceptual}), including a strong decline in the fraction of plant
that is leaf declines with increasing size (Fig. \ref{f-mass_fraction}),
and increasing reproductive allocation. In contrast, basal-area growth
continues to increase with size \citep{Sillett-2010, Stephenson-2014},
due to an increasing influence of stem
turnover. Diameter growth shows a weakly hump shaped
curve \citep{Herault-2011}, tapering off slightly at larger
sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
mainly because of increased reproductive allocation in older trees
(Fig?). All growth measures decrease sharply with size when expressed as
relative growth rates \citep{Iida-2014}.

@@

@@

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
changes with plant size ( \ref{f-lmA_growth_size}).
Decreasing $\phi$ has two impacts on growth rate. First, lower
$\phi$ increases marginal leaf deployment
(${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on
construction costs. Second, lower $\phi$ decreases net production
(${\rm d}P / {\rm d}t$), due to increased leaf turnover. Whether
lower $\phi$ increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing $\phi$ increases growth
rate. When plants are large, the influence of $\phi$ on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{f-mass_fraction}). The net result is that at larger sizes, low
$\phi$ is no longer advantageous for growth (Fig.
\ref{f-lmA_growth_size}d).

@@

Two other important predictions arise directly from linking traits to
growth rate. First, traits moderate the
responsiveness of growth to changes in light environment. This response
arises simply because species with higher potential growth rate have
greater potential plasticity. Recent analyses have shown that species
with low $rho$ increase growth more substantially with incraeses in
light \citep{Ruger-2012}. This is exactly the behaviour
exhibited in our model (Fig. \ref{f-growth_light_dia}), at least for small
to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
of low $\phi$ are lost (Fig. \ref{f-growth_light_dia}). This results also
agrees with empirical results showing that growth for species with low
$rho$ actually decreases as size
increases \citep{Ruger-2012}. In contrast, growth rate tends
to increase in size for the species with highest
$\rho$ \citep{Ruger-2012}.

@@

\subsection{Growth is more than just photosynthesis}\label{tree-growth-is-more-than-just-photosynthesis}

@@

\subsection{Shade tolerance}\label{shade-tolerance}

Second, species with low $\phi$ and low $\rho$ are predicted to be
less shade tolerant (Fig. \ref{f-wplcp}). At low $\phi$ ($\rho$),
leaf (sapwood) turnover is higher and thus a greater light income is
needed to offset these costs. As previously suggested
 \citep{Givnish-1988}, shade-tolerance also decreases with
height because as size increase, the total amount of energy needed to
offset respiratory and turnover costs in the stem also increases. These
finding matches well known empirical patterns, where both $\phi$ and
$\rho$ have been linked to shade tolerance \citep{Poorter-2006, Lusk-2008,Osunkoya-1996}.

@@

\subsection{Implications for trait-based approaches}\label{implications-for-trait-based-approaches}

@@

\section{Acknowledgments}

@@

@@

\section{Tables}\label{tables}

\begin{table}[ht]
 \caption{Key variables
%  $a = l + ss + sb +  r$,
%  $M_{sh}$,
%  $M_t = M_a + M_{sh}$,
%  $st = stem total$
}
\centering

@@

  $b$   & kg  & biomass originating from parent plant\\
  $M_i$ & kg  & mass retained on plant in tissue $i$\\
  $A_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
  $h$   & m  & height of plant\\
  $D$   & m  & diameter of plant stem\\

  $y$ & kg kg$^{-1}$ & Yield; ratio of carbon fixed in mass per carbon assimilated \\
  $p$ & kg yr$^{-1}$ m$^{-2}$  & photosynthesic rate per unit area \\
  $r_i$ & kg yr$^{-1}$ kg$^{-1}$  & respiration per unit tissue mass \\
  $k_i$ & yr$^{-1}$ & turnover rate for tissue \\
  $A$ & kg yr$^{-1}$ & total plant photosynthesis \\
  $R$ & kg yr$^{-1}$ & total plant respiration \\
  $T$ & kg yr$^{-1}$ & total plant turnover \\
  \hline
  \end{tabular}
\end{doublespace}}
\label{tab:definitions}
\end{table}

@@

{\footnotesize  % smaller font, double space
 \begin{doublespace}
  \begin{tabular}{p{2.5cm}p{3.5cm}p{5cm}p{4cm} }\\ \hline
  Variable & Function & Marginal cost & Growth rate\\ \hline
  \multicolumn{4}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
  height &
    $h = \alpha_1 \, A_{\rm l}^{\beta_1}$ &
    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} =  \beta_1 \alpha_1 A_{\rm l}^{\beta_1-1}$ &
    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
  sapwoood area &
    $A_\textrm{ss} = \theta \, A_{\rm l}$ &
    $\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} = \theta$ &
    $\frac{{\rm d}A_\textrm{ss}}{{\rm d}t}  =\frac{{\rm d}A_\textrm{ss}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
  bark area &
    $A_{\rm sb} = b \, \theta \, A_{\rm l}$ &
    $\frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} = b \, \theta$ &
    $\frac{{\rm d}A_{\rm sb}}{{\rm d}t} = \frac{{\rm d}A_{\rm sb}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\  \\
  \multicolumn{4}{l}{\textbf{b) Derived equation for mass of tissue }} \\
  leaf mass &
    $M_{\rm l} = \phi \, A_{\rm l} $ &
    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
  sapwood mass &
    $M_\textrm{ss} = \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( h + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
    $\frac{{\rm d}M_\textrm{ss}}{{\rm d}t}  = \frac{{\rm d}M_\textrm{ss}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
  bark mass &
    $M_{\rm sb} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
    $\frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( h + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
    $\frac{{\rm d}M_{\rm sb}}{{\rm d}t}  = \frac{{\rm d}M_{\rm sb}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\
  root mass &
    $M_{\rm r} = \alpha_3 \, A_{\rm l}$ &
    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_3$  &
    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ \\ \\

>>>>

% Enable equation numbers in Table
% Following http://tex.stackexchange.com/questions/62974/numbering-equations-in-tabular-environment
\newcommand{\eqncounter}[1]{%
\mbox{}\refstepcounter{equation}%
$(\theequation)$%
\label{#1}
}

@@

% Enable cross referencing to Evidence document
\usepackage{xr}
\externaldocument{ms-supp}
\externaldocument{figure_conceptual}

@@

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
% \doublespacing
% \linenumbers
\section{Summary}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item We ...
\end{enumerate}

% Consider Ecol Letters I deas and Perspectives: maximum of 7500 words (main text), and 10 figures, tables or boxes.

\section{To do}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item New refs: \citep{Ishihara-2016, Duursma-2016, Gibert-2016, Falster-2016, Kunstler-2015,Falster-2015b, Falster-2015a,Wenk-2015,Li-2014}
\item Table near intro summarising empirical patterns to be explained by integrated model?
\end{enumerate}

\section{Introduction}

@@

\section{Conceptual framework for understanding height, basal area and diameter growth}

@@

Equations \ref{eq:dbdt} -\ref{eq:dDdt} shown in Fig. \ref{fig:conceptual}.
\setcounter{equation}{6}

Requires definitions from Table \ref{tab:definitions}.

(check this material covered in table: Here, $M, r$, and $k$ refer to the mass, respiration rate, and
turnover rate of different tissues, denoted by subscripts $l$=leaves,
$b$=bark, $s$=sapwood and $r$=roots. $A$ is the assimilation
rate of CO$_2$ per leaf area and $y$ is yield: the fraction of
assimilated carbon fixed in biomass (the remaining fraction being lost
as growth respiration) (see Table \ref{tab:params} for units and
definitions). Photosynthesis is proportional to leaf area,
$A_{\rm l} = M_{\rm l} / \phi$, where $\phi$ is leaf mass per area.
The total mass of living tissues $M_{\rm a}=M_{\rm l}+M_{\rm b}+M_{\rm s}+M_{\rm r}.$
)

@@

\begin{equation}\label{eq:dbdt_full}
\underbrace{\strut \frac{{\rm d}B}{{\rm d}t}}_\textrm{net biomass production}
  = \underbrace{\strut y}_\textrm{yield}
    \big( \underbrace{\strut A_{\rm l} \, p}_\textrm{photosynthesis} -
     \underbrace{\strut \,\sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, r_\textrm{i}}}_\textrm{respiration}\big)
    - \underbrace{\strut \sum_\textrm{i=l,b,s,r}{M_\textrm{i} \, k_\textrm{i}}}_\textrm{turnover}.
\end{equation}

The only potentially controversial element of in eq. \ref{eq:dbdt} \ref{eq:dbdt_full} is that it assumes tissues lost via turnover are replaced before energy is allocation to either new growth or reproduction \citep{Thornley-2000}. This assumption is likely to hold true for most woody plants, because the ..... However, the assumption may not hold for some herbs or annuals, where the swicth to reproduction often entails run-down in the vegetative part of the plant (refs XX) \citep{ Thornley-2000}.

To model growth in either height, basal area or stem diameter requires that account not just for mass production, but also for the costs of building new tissues, allocation to reproduction, and architectural layout. Mathematically, height growth can be decomposed into a product of physiologically relevant terms \citep{Falster-2011, Gibert-2016} (Figure \ref{fig:conceptual})

@@

@@

In a similar way, basal area ($A_{\rm st}$) increment can be
expressed as the sum of increments in sapwood, bark \& heartwood areas
($A_{\rm s}, A_{\rm b}, A_{\rm h}$ respectively):
$\frac{{\rm d}A_{\rm st}}{{\rm d}t}= \frac{{\rm d}A_{\rm b}}{{\rm d}t} + \frac{{\rm d}A_{\rm s}}{{\rm d}t} + \frac{{\rm d}A_{\rm h}}{{\rm d}t}$.
We now have an equation for basal area growth that contains many of the same
elements as eq. \ref{eq:dhdt}, see eq. \ref{eq:dast}.

Diameter growth is then given by the geometric relationship between stem
diameter ($D$) and $A_{\rm st}$, see eq. \ref{eq:dDdt}

@@

To address the first of these challenges, we use a functional balance
model \citep{Yokozawa-1995, Falster-2011} assuming
1) an allometric relation between total leaf area and plant height, and
2) a constant ratio between total leaf area and each of sapwood
cross-sectional area, bark cross-sectional area, and total root mass.
These assumptions, which are well-supported by data (Fig.
\ref{fig:assumptions}), lead immediately to an allometric model for plant
size (Table \ref{tab:allometry}, see Supplementary material for
derivations). Substituting from Table \ref{tab:allometry} into eqs.
\ref{eq:dhdt}, \ref{eq:dast}, \ref{eq:dDdt} then gives an explicit model
for plant growth in relation to size.

@@

\section{Derivation of a simple allometric model of plant function}

Based on empirically observed allometries (see main text), we assume an
allometric log-log scaling relationship between the accumulated leaf
area of a plant and its height: Eq. \ref{eq:ha}.

We follow the model of \citep{Yokozawa-1995} describing the
vertical distribution of leaf area within the crowns of individual
plants. This model can account for a variety of canopy profiles through
a single parameter $\eta$. Setting $\eta=1$ results in a conical
canopy, as seen in many conifers, while higher values, e.g. $\eta=12$
, give a top-weighted canopy profile similar to those seen among
angiosperms. We also assume that each unit of sapwood area supports a fixed area of
leaf (the pipe model, \citep{Shinozaki-1964}), so that the total
canopy area of a plant relates to basal sapwood area. This leads to a solution for the total mass of sapwood in the plant as a function leaf area and height (Eq. \ref{eq:ms2}).

Bark and phloem tissue are modelled using an analogue of the pipe model,
leading to a similar equation as that for sapwood mass (eq.
\ref{eq:ms2}). Cross sectional-area of bark per unit leaf area is
assumed to be a constant fraction $b$ of sapwood area per unit leaf
area such that (Eq. \ref{eq:mb}).

Also consistent with pipe-model assumption, we assume a fixed ratio of
root mass per unit leaf area (Eq. \ref{eq:mr}).

\subsubsection{Embedding trait-related trade-offs}

@@

As for leaf, cheaper stem construction (lower $\rho$) decreases the
cost of deploying a unit of leaf area, and may thereby increase growth
rates. However, the cost of cheap stem construction must also be
accounted for. However, the direct physiological trade-offs of lower
$\rho$ are less understood than for $\phi$. One possibility is that
lower $\rho$ decreases mechanical strength resulting in higher
mortality \citep{Chave-2009, Wright-2010}. Under
this scenario, decreasing $\rho$ will always provide a growth
advantage, because the costs of low $\rho$ are not realised within the
terms of eqs. \ref{eq:dhdt} and \ref{eq:dast} (Table
\ref{tab:trade-offs}). In addition, lower $\rho$ may increase turnover
of sapwood. Mirroring our assumptions for leaves, we let sapwood
turnover decrease with $\rho$,
i.e.$k_{\rm s}=\alpha_5 \, \rho^{-\beta_5}$, with $\beta_5 > 1$.

@@

\subsection{Analysis}

\newpage
\section{Results and Discussion}

@@

This new growth model captures the intrinsically size-dependent nature
of plant growth, recovering well-known
empirical patterns \citep{Sillett-2010, King-2011}. Height growth shows a hump-shaped pattern
with size, first increasing then decreasing. This pattern results from
systematic changes in the four components of eq. \ref{eq:dhdt} with size
(Fig. \ref{fig:conceptual}), including a strong decline in the fraction of plant
that is leaf declines with increasing size (Fig. \ref{fig:mass_fraction}),
and increasing reproductive allocation. In contrast, basal-area growth
continues to increase with size \citep{Sillett-2010, Stephenson-2014},
due to an increasing influence of stem
turnover. Diameter growth shows a weakly hump shaped
curve \citep{Herault-2011}, tapering off slightly at larger
sizes, in part because of allometric effects (eq. \ref{eq:dDdt}, but
mainly because of increased reproductive allocation in older trees
(Fig?). All growth measures decrease sharply with size when expressed as
relative growth rates \citep{Iida-2014}.

@@

For growth traits can derive rule:
\begin{equation}\label{eq:dg4}
\frac{
\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
{\frac{ {\rm d}B} {{\rm d}t}}
<
\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
\end{equation}

We can also solve for trait values $x$ maximising growth rate, i.e. the value
giving $\partial G /\partial x = 0$:
\begin{equation}\label{eq:max}
\frac{
\frac{\partial \left( \frac{ {\rm d}B} {{\rm d}t}\right)}{\partial x} }
{\frac{ {\rm d}B} {{\rm d}t}}
=
\frac{ \frac{\partial \left(\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}\right)
}{\partial x}}{\frac{{\rm d}M_{\rm a}} {{\rm d}A_{\rm l}}}.
\end{equation}

@@

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
changes with plant size ( \ref{fig:lmA_growth_size}).
Decreasing $\phi$ has two impacts on growth rate. First, lower
$\phi$ increases marginal leaf deployment
(${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$ by economising on
construction costs. Second, lower $\phi$ decreases net production
(${\rm d}P / {\rm d}t$), due to increased leaf turnover. Whether
lower $\phi$ increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing $\phi$ increases growth
rate. When plants are large, the influence of $\phi$ on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{fig:mass_fraction}). The net result is that at larger sizes, low
$\phi$ is no longer advantageous for growth (Fig.
\ref{fig:lmA_growth_size}d).

\begin{equation} \label{eq:G6}
\frac{(B_4-1) A_{\rm l} A_4\phi ^{-B_4}}{c_2 - A_{\rm l} A_4 \phi ^{1-B_4}}
< \frac{1}{\phi
 + \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_\textrm
 {b}}{{\rm d}A_{\rm l}} + \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}}.
\end{equation}

@@

Two other important predictions arise directly from linking traits to
growth rate. First, traits moderate the
responsiveness of growth to changes in light environment. This response
arises simply because species with higher potential growth rate have
greater potential plasticity. Recent analyses have shown that species
with low $rho$ increase growth more substantially with incraeses in
light \citep{Ruger-2012}. This is exactly the behaviour
exhibited in our model (Fig. \ref{fig:growth_light_dia}), at least for small
to intermediate sizes. At larger sizes (D=0.2m) however, the advantages
of low $\phi$ are lost (Fig. \ref{fig:growth_light_dia}). This results also
agrees with empirical results showing that growth for species with low
$rho$ actually decreases as size
increases \citep{Ruger-2012}. In contrast, growth rate tends
to increase in size for the species with highest
$\rho$ \citep{Ruger-2012}.

@@

\subsection{Growth is more than just photosynthesis}

@@

\subsection{Shade tolerance}

Second, species with low $\phi$ and low $\rho$ are predicted to be
less shade tolerant (Fig. \ref{fig:wplcp}). At low $\phi$ ($\rho$),
leaf (sapwood) turnover is higher and thus a greater light income is
needed to offset these costs. As previously suggested
 \citep{Givnish-1988}, shade-tolerance also decreases with
height because as size increase, the total amount of energy needed to
offset respiratory and turnover costs in the stem also increases. These
finding matches well known empirical patterns, where both $\phi$ and
$\rho$ have been linked to shade tolerance \citep{Poorter-2006, Lusk-2008,Osunkoya-1996}.

@@

\subsection{Implications for trait-based approaches}

@@

\section{Acknowledgements}

@@

\section{Tables}

\begin{table}[ht]
 \caption{Key phenomena to be explained}
\centering

{\footnotesize  % smaller font, double space
 \begin{doublespace}
  \begin{tabular}{p{7cm}p{4cm}}
  \hline
  Phenomena & References \\
  \hline
  hump shape growth & \citep{King-2005} \\
  \hline
  \end{tabular}
\end{doublespace}}
\label{tab:phenomena}
\end{table}

@@

\begin{table}[ht]
 \caption{Key variables}
\centering

@@

  $B$   & kg  & biomass originating from parent plant\\
  $M_i$ & kg  & mass retained on plant in tissue $i$\\
  $A_i$ & m$^2$  & surface-area or cross-sectional-area of tissue $i$\\
  $H$   & m  & height of plant\\
  $D$   & m  & diameter of plant stem\\
  $\alpha_y$ & kg kg$^{-1}$ & Yield; ratio of carbon fixed in mass per carbon assimilated \\
  $P$ & kg yr$^{-1}$ & total plant photosynthesis \\
  $R$ & kg yr$^{-1}$ & total plant respiration \\
  $T$ & kg yr$^{-1}$ & total plant turnover \\
  $p$ & kg yr$^{-1}$ m$^{-2}$  & photosynthetic rate per unit area \\
  $r_i$ & kg yr$^{-1}$ kg$^{-1}$  & respiration per unit tissue mass \\
  $k_i$ & yr$^{-1}$ & turnover rate for tissue \\
  \hline
  \end{tabular}
\end{doublespace}}
\label{tab:definitions}
\end{table}

@@

{\footnotesize  % smaller font, double space
 \begin{doublespace}
  \begin{tabular}{p{2.5cm}p{3cm}p{4cm}p{2.5cm}p{0.75cm}}\\ \hline
  Variable & Function & Marginal cost & Growth rate & Eq.\\ \hline
  \multicolumn{5}{l}{\textbf{a) Assumed relationship to leaf area}} \\ \\
  height &
    $H = \alpha_1 \, A_{\rm l}^{\beta_1}$ &
    $\frac{{\rm d}H}{{\rm d}A_{\rm l}} =  \beta_1 \alpha_1 A_{\rm l}^{\beta_1-1}$ &
    $\frac{{\rm d}H}{{\rm d}t}  = \frac{{\rm d}H}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ &  \eqncounter{eq:ha}\\
  sapwood area &
    $A_{\rm s} = \theta \, A_{\rm l}$ &
    $\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} = \theta$ &
    $\frac{{\rm d}A_{\rm s}}{{\rm d}t}  =\frac{{\rm d}A_{\rm s}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:a}\\
  bark area &
    $A_{\rm b} = b \, \theta \, A_{\rm l}$ &
    $\frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} = b \, \theta$ &
    $\frac{{\rm d}A_{\rm b}}{{\rm d}t} = \frac{{\rm d}A_{\rm b}}{{\rm d} A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:a}\\  \\
  \multicolumn{5}{l}{\textbf{b) Derived equation for mass of tissue }} \\
  leaf mass &
    $M_{\rm l} = \phi \, A_{\rm l} $ &
    $\frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}} = \phi$ &
    $\frac{{\rm d}M_{\rm l}}{{\rm d}t}  = \frac{{\rm d}M_{\rm l}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ml}\\
  sapwood mass &
    $M_{\rm s} = \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
    $\frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} = \theta\, \rho\, \eta_c\, \big( h + A_{\rm l}\, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
    $\frac{{\rm d}M_{\rm s}}{{\rm d}t}  = \frac{{\rm d}M_{\rm s}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:ms2}\\
  bark mass &
    $M_{\rm b} = b\, \theta \, \rho \, \eta_c \, A_{\rm l} \, h $ &
    $\frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} = b \, \theta \, \rho \, \eta_c\big( h + A_{\rm l} \, \frac{{\rm d}H}{{\rm d}A_{\rm l}} \big)$ &
    $\frac{{\rm d}M_{\rm b}}{{\rm d}t}  = \frac{{\rm d}M_{\rm b}}{{\rm d}A_{\rm l}} \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mb}\\
  root mass &
    $M_{\rm r} = \alpha_3 \, A_{\rm l}$ &
    $\frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}} = \alpha_3$  &
    $\frac{{\rm d}M_{\rm r}}{{\rm d}t}  = \frac{{\rm d}M_{\rm r}}{{\rm d}A_{\rm l}}  \, \frac{{\rm d}A_{\rm l}}{{\rm d}t}$ & \eqncounter{eq:mr}\\ \\
